Molecular transport in human pial perivasculature

### 2024-11-20 paper/supplementary_main.tex
Simplify notation.
Former-commit-id: 97a165d7f0321b6c335f5b019593e033c356b3ef

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s (\hat c) ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces. To
complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{T}$, write $\hat{c}^i = \hat{c}|_{\Lambda^i}$, and let
$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
concentrations, we set that the concentrations when viewed from each
edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
\end{equation}
Moreover, to ensure \discuss{mass conservation}, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n_i(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n_i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n_i(y_{\rm in}) = 1, \quad
  n_i(y_{\rm{out}}) = -1, \quad
  n_i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
\end{equation}

%% For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n_i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
%% $$
%% The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$. For any $\mathsf{v} \in \mathcal{B}$, we have 
%% \begin{equation}
%% \sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0  \label{eq:flux_conservation}
%% \end{equation}

@@

On the interface between the brain and the CSF spaces
$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
following interface conditions, which represent a semi--permeable
interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
%% \begin{subequations}
%% \begin{alignat}{2}
%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
%% \end{alignat} 
%% \end{subequations}
\begin{subequations}
  \begin{alignat}{2}
  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
\end{alignat} 
  \label{eq:interface:3d}%
\end{subequations}%
where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
over the interface, outward-pointing when viewed from $\Omega_{\rm
  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
constant.

@@

We supplement~\eqref{eq:3d_pde} and~\eqref{eq:interface:3d} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
\begin{subequations}
\begin{alignat}{2}
(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
\end{alignat} 
\end{subequations}
where again $\bm n$ denotes the outward-pointing boundary normal.

@@

For the perivascular networks, we assume no molecular efflux at all
end nodes $\mathcal{T}$ via a homogeneous Neumann condition
augmenting~\eqref{eq:1d_pde}:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{T}.
\end{equation}

@@

\subsubsection{Flow in perivascular networks induced by pressure gradients}

To model steady flow induced by pressure differences between end nodes
in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
internal nodes $\mathcal{B}$ and end nodes $\mathcal{T}$, we consider
the following system of hydraulic network
equations~\cite{tithof2022network, daversin2022geometrically,
  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
\rightarrow \R$, which represent the fluid flux and the average
pressure across cross-sections of the PVSs, respectively. These are
defined for each $\Lambda^i$ by solving:
\begin{subequations}
\begin{alignat}{2}
  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
  \quad \text{ in } \Lambda^i , \\ 
  - \partial_s \hat{q} & = 0 &&
  \quad \text{ in } \Lambda^i ,  
\end{alignat} 
\label{eq:1d_flow}%
\end{subequations}%
where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}
The PVS flux $\hat{q}$ and the associated average PVS velocity in the
axial direction $\hat{u}$ are directly related by 
area
\begin{equation}
  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
\end{equation}

In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
the fluid pressure and conservation of flux at each internal node. Write
$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
\begin{subequations}
\begin{align}
  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
\end{align}
\end{subequations}
Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
network end nodes:
\begin{equation}
  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{T} .
\end{equation}
In the case where modelling flow in the perivascular networks
$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
fluid pressures $p_0$.

\mer{MER: Stopped here.}

@@

Here we provide details on the numerical method used to solve the coupled system. We are given a mesh $\mathcal{T}_h $, consisting of elements denoted by $E$, of $\Omega= \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$.  The collection of all interior facets to $\Omega_{\rm PAR}$ (resp.  to $\Omega_{\mathrm{CSF}}$) is  denoted by $\mathcal{F}_{h,b}$ (resp. $\mathcal{F}_{h,c}$).  The union of all interior facets $\mathcal{F}_i = \mathcal{F}_{h,b} \cup \mathcal{F}_{h,c}$. Facets lying on the pial interface are denoted by $\mathcal{F}_{\Gamma_{\rm pia}}$, on the skull by $\mathcal{F}_{\Gamma_{\rm skull}}$, and on the spinal canal by $\mathcal{F}_{\Gamma_{\rm SAS}}$. Note that $\mathcal{F}_i \cap \mathcal{F}_{\Gamma_{\rm pia}} = \emptyset $. The union of all the facets is denoted by $\mathcal{F}_h$. The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is denoted by $\mathcal{E}_{h}$.

\subsubsection{Solute transport}
For the 3D concentration, we use broken polynomial spaces and interior penalty discontinuous Galerkin (DG) methods with upwinding in order to capture the sharp boundary layers.  For DG, we utilize the discrete space 
\begin{equation}
V_h^k = \{v_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
\end{equation}
To discretize the diffusion term, we utilize symmetric weighted interior penalty dG bilinear forms. We refer to \cite{ern2009discontinuous} and \cite[Section 4.5.2.3]{di2011mathematical} for details on this method.  We define 
\begin{multline}
a_h (c_h, v_h) = \sum_{E \in \mathcal{T}_h } \int_{E} D \nabla (\phi c_h) \cdot \nabla v_h + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c_h][v_h] 
- \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c_h \}_w \cdot \bm n_F [v_h]  + \{D \phi\nabla v_h \}_w \cdot \bm n_F [c_h] \right).
 \nonumber
 \end{multline}
In the above, for $F\in \mathcal{F}_h^i$ with a normal $n_F$ pointing from $E_F^1$ to $E_F^2$,  the jump is given by $[v_h] = v_h{}_{\vert_{E_F^1}} - v_h{}_{\vert_{E_{F}^2}}$, and the weighed average is defined as follows.
\begin{equation}
\{v_h\}_{w} = \frac{ \kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^1}} + \frac{\kappa_{\vert_{E_F^1}}}{\kappa_{ \vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}} v_h{}_{\vert_{E_{F}^2}}, \quad \text{where} \quad \kappa = D \phi.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the diffusion coefficient given by 
$$
\gamma_{D,F} = \frac{ \kappa_{\vert_{E_F^1}}\kappa_{\vert_{E_F^2}}}{\kappa_{\vert_{E_F^1}}+ \kappa_{\vert_{E_F^2}}},
$$
and $\eta$ is a user specified penalty to be determined later.

For the convection term, we use upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the references therein. Define 
\begin{equation*}
a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ] 
\end{equation*}
The discrete formulation for \eqref{eq:3d_pde} then reads 
\begin{multline}
\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\Gamma_{\rm SAS}} \gin v_h , \quad \forall v_h \in V_h^k.
\end{multline}
For approximating the solution of the 1D equations in the network, we use the space 
\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
\}, 
\]
where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,

>>>>

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s (\hat c) ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces. To
complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{T}$, write $\hat{c}^i = \hat{c}|_{\Lambda^i}$, and let
$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
concentrations, we set that the concentrations when viewed from each
edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
  \label{eq:conc_continuity}
\end{equation}
Moreover, to ensure \discuss{mass conservation}, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n_i(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n^i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n^i(y_{\rm in}) = 1, \quad
  n^i(y_{\rm{out}}) = -1, \quad
  n^i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
  \label{eq:def:sign}
\end{equation}

%% For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n^i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
%% $$
%% The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$. For any $\mathsf{v} \in \mathcal{B}$, we have 
%% \begin{equation}
%% \sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n^i(\mathsf{v}) = 0  \label{eq:flux_conservation}
%% \end{equation}

@@

On the interface between the brain and the CSF spaces
$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
following interface conditions, which represent a semi--permeable
interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
%% \begin{subequations}
%% \begin{alignat}{2}
%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
%% \end{alignat} 
%% \end{subequations}
\begin{subequations}
  \begin{alignat}{2}
  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
\end{alignat} 
  \label{eq:3d_interface}%
\end{subequations}%
where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
over the interface, outward-pointing when viewed from $\Omega_{\rm
  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
constant.

@@

We supplement~\eqref{eq:3d_pde} and~\eqref{eq:3d_interface} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
\begin{subequations}
\begin{alignat}{2}
(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \cup \Gamma_{\rm SC}  .
\end{alignat} 
\label{eq:3d_bcs}
\end{subequations}
where again $\bm n$ denotes the outward-pointing boundary normal.

@@

For the perivascular networks, we assume no molecular efflux at all
end nodes $\mathcal{T}$ via a homogeneous Neumann condition
augmenting~\eqref{eq:1d_pde}:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{T}.
    \label{eq:1d_bcs}
\end{equation}

@@

\subsubsection{Steady flow in perivascular networks -- induced by pressure differences}

To model steady flow induced by pressure differences between end nodes
in a perivascular network $\Lambda$, with edges $\{\Lambda^i\}_i$,
internal nodes $\mathcal{B}$ and end nodes $\mathcal{T}$, we consider
the following system of hydraulic network
equations~\cite{tithof2022network, daversin2022geometrically,
  gjerde2024directional}. The unknowns are the PVS flux $\hat{q} :
\Lambda \rightarrow \R$ and pressure $\hat{p} : \Lambda
\rightarrow \R$, which represent the fluid flux and the average
pressure across cross-sections of the PVSs, respectively. These are
defined for each $\Lambda^i$ by solving:
\begin{subequations}
\begin{alignat}{2}
  \hat{q} + \frac{A \kappa}{\mu} \partial_{s} \hat{p} & = 0 &&
  \quad \text{ in } \Lambda^i , \\ 
  - \partial_s \hat{q} & = 0 &&
  \quad \text{ in } \Lambda^i ,  
\end{alignat} 
\label{eq:1d_flow}%
\end{subequations}%
where $A$ is the area of the PVS cross-sections, $\mu$ is the CSF viscosity as before, and $\kappa$ is derived from an assumption of Poiseuille flow in the annular cross-section of the PVS as~\cite{tithof2022network}: 
\begin{equation}
\kappa = \frac18 \left( R_2^2 + R_1^2 - \frac{1}{\ln(R_2/R_1)} (R_2^2- R_1^2) \right). 
\end{equation}
The PVS flux $\hat{q}$ and the associated average PVS velocity in the
axial direction $\hat{u}$ are directly related by the cross-sectional area
\begin{equation}
  \hat{q}^i = A^i \hat{u}^i \qquad \foralls \Lambda^i.
\end{equation}

In addition, to complete~\eqref{eq:1d_flow}, we impose continuity of
the fluid pressure and conservation of flux at each internal node. Write
$\hat{p}^i = \hat{p}|_{\Lambda^i}$. These conditions then read as: for
each $y \in \mathcal{B}$ with connected edges $\mathcal{E}(y)$:
\begin{subequations}
\begin{align}
  \hat{p}^i(y) &= \hat{p}^j(y) \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y), \\
  \sum_{\Lambda^i \in \mathcal{E}(y)} \hat{q}^i n^i(y) &= 0  .
\end{align}
\end{subequations}
where $n^i$ is as defined by~\eqref{eq:def:sign}. Finally, to drive this flow, we impose given fluid pressures $p_0$ at the
network end nodes:
\begin{equation}
  \hat{p}(y) = p_0(y) \qquad y \in \mathcal{T} .
\end{equation}
In the case where modelling flow in the perivascular networks
$\Lambda_a, \Lambda_v$ induced by CSF production, we impose the fluid
pressure $p_{\rm CSF}$ computed in the CSF spaces $\Omega_{\rm CSF}$
(and its harmonic extensions into $\Omega_{\rm PAR}$) as the given
fluid pressures $p_0$.

@@

We solve the models presented in~\Cref{sec:more_maths} numerically
using finite element methods. The following subsections provide more
details, model-by-model.

\subsubsection{Computational mesh entities: definitions and notation}

We consider a mesh $\mathcal{T} = \{ E \}$ of $\Omega = \Omega_{\rm
  PAR} \cup \Omega_{\rm CSF}$, consisting of tetrahedral mesh cells
$E$, and conforming to the domains $\Omega_{\rm PAR}$ and $\Omega_{\rm
  CSF}$ and to the CSF-brain interface $\Gamma_{\rm pia} \cup
\Gamma_{\rm LV}$. We denote the restriction of $\mathcal{T}$ to
$\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ by $\mathcal{T}_{\rm PAR}$
and $\mathcal{T}_{\rm CSF}$, respectively. The collection of all
interior facets (faces of the mesh cells) in $\mathcal{T}_{\rm PAR}$
(resp. in $\mathcal{T}_{\rm CSF}$) is denoted by $\mathcal{F}_{i, \rm
  PAR}$ (resp. $\mathcal{F}_{i, \rm CSF}$). We define the union of
facets interior both to $\mathcal{T}_{i, \rm PAR}$ and
$\mathcal{T}_{i, \rm CSF}$ as $\mathcal{F}_i = \mathcal{F}_{i, \rm
  PAR} \cup \mathcal{F}_{i, \rm CSF}$. Facets lying on the pial
interface $\Gamma_{\rm pia}$ are denoted by $\mathcal{F}_{\rm pia}$,
on the outer (arachnoid) boundary $\Gamma_{\rm AM}$ by
$\mathcal{F}_{\rm AM}$, on the interface toward the spinal cord
$\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$ and on the interface
towards the spinal SAS $\Gamma_{\rm SSAS}$ by $\mathcal{F}_{\rm
  SSAS}$. The union of all facets in $\mathcal{T}$ is denoted by
$\mathcal{F}$.

%Note in particular that $\mathcal{F}_i \cap \mathcal{F}_{pia}} =
%  \emptyset$.

%The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is
%denoted by $\mathcal{E}_{h}$.

\subsubsection{Finite element solution of coupled 3D-1D solute transport equations }

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time and
a discontinuous Galerkin (DG) finite element method with upwinding in
space to capture the sharp boundary layers.

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h^k = \{c \in L^2(\discuss{\mathcal{T}}): \;  c \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\mathcal{T})$ is the space of square-integrable functions
defined relative to $\mathcal{T}$ and $\mathbb{P}^k(E)$ denotes the
space of polynomials of degree $\leqslant k$ defined over the
tetrahedra $E$. To discretize the diffusion term in~\eqref{eq:3d_pde},
we use a symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. \discuss{Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.}
Define
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D \discuss{\phi} \nabla c \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$ with facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$, the jump $[\cdot]$ is given by $[v] =
v|_{E^1} - v|_{E^2}$, and the weighed average $\{ \cdot \}_w$ is
defined as:
\begin{equation}
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2},
  \quad \text{where} \quad \kappa_i = \phi D|_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the \discuss{porosity-weighted} diffusion coefficient given by 
\begin{equation}
  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter.

\mer{MER. Stopped here.}

For the convection term, we use upwinding, see \cite[Section
  2.3.1]{di2011mathematical} and the references therein. For discrete
functions $c, v \in V_h^k$, define
\begin{equation*}
a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ] 
\end{equation*}
The discrete formulation for \eqref{eq:3d_pde} then reads 
\begin{multline}
\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\Gamma_{\rm SAS}} \gin v_h , \quad \forall v_h \in V_h^k.
\end{multline}
For approximating the solution of the 1D equations in the network, we use the space 
\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
\}, 
\]
where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,
